---
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#mnarmple: A package for calculating the MNAR MPLE likelihood for the cause specific Cox model with missing failure type

The `mnarmple` package calculates the negative log likelihood of the MNAR MPLE likelihood equation for a given set of data and parameters, and can be used with tho `optim()` function to get parameter estimates based on maximizing the log-likelhood. It also includes a function that can be used with the `boot()` function from the package `boot`.

##Installation
```{install}
devtools::install_github("blangworthy/mnarmple")
library(mnarmple)
```

##Examples

```{r example, results="hide", message=FALSE}
library(gtools)
library(data.table)
library(splines2)
library(fastDummies)
#devtools::install_github("blangworthy/mnarmple")
library(mnarmple)
#####Example without competing risks
###In this example we simulate 4 non-competing events and calculate the martingale and counting process
###covariance and correlation matrices. All events have exponential 1 distribution and we estimate the
###covariance/correlation matrices at timepoint 1.

n <- 5000 # sample size
###Single binary covariate
x <- rbinom(n,1,0.5)
ncause <- 2

###True Cox model log hazard ratios
b11 <- 0
b21 <- 0.69
###Cause specific failure time
t1 <- rweibull(n,shape = 1,scale = exp(-b11/1*x))
t2 <- rweibull(n,shape = 1.5,scale = exp(-b21/1.5*x))


tall <- cbind(t1,t2)
####minimum time and cause indicator
t <- apply(tall,1,min)
cause <- apply(tall,1,which.min)

####censoring time
c <- rbeta(n,shape1=0.1,shape2 = 3)*5

###observied time and censoring indicator
combined <- cbind(t,c)
z <- apply(combined,1,min)
cens <- apply(combined,1,function(x)ifelse(which.min(x)==1,1,0))

###True missingness model coefficients
ax1 <- 0.5
ay1 <- -0.5
ay2 <- 0.5
###Indicator for if cause is observed based on logit model (MNAR if ay1 doesn't equal ay2)
prob <- x*ax1 + as.numeric(cause==1)*ay1 + as.numeric(cause==2)*ay2
o <- rbinom(n,1,gtools::inv.logit(prob))


###Basis function to estimate baseline hazard ratio as cubic spline
basis <- splines2::bSpline(z,knots =quantile(z,c(1/3,2/3)),degree = 3 ,intercept = T)
splinevars <- paste0("bas",1:ncol(basis))
colnames(basis) <- splinevars

###Create 'observed' data
estdat <- data.table(x,z,cause,cens,o,basis)
estdat$cause <- estdat$cause *estdat$o

###Supplied starting parameters
paramszero <- c(0,0,0,rep(1,ncol(basis)))
names(paramszero) <- c("ax1","b11","b21",paste0("bas2",1:ncol(basis)))

estszero <- optim(paramszero,method = "Nelder-Mead",fn = likesplinefunctionmnarfixed,dat = estdat,mnarparams=c(0.5,-0.5),timevar = "z",xvarsmiss="x",xvarscox="x",splinevars=splinevars,ncause=ncause)

```
